Regional-scale Base-flow Index of Gauged and Ungauged Basins in Drylands

### 2024-11-26 index.qmd
New

: **Table 1.** Basin-characteristic variables used as predictors in XGBoost model. {#tbl-predictors}

@@

The long-term average BFI for the 245 gauged reaches across Arizona is illustrated in @fig-instrumented-bfi.The highest BFI values (\>0.9) are found along the Grand Canyon in northwestern Arizona. The highly karstic geology of this region facilitates the rapid movement of subsurface flow to surface water and spring outlets [@chambless_deep-karst_2023]. Relatively high BFI values (\>0.8) are found at the spring-fed headwaters of the Verde River (Del Rio Spring) and the spring-fed headwaters of Fossil Creek. These results are consistent with interpolated BFI values reported by @wolock2003.

@@

@@

: Table 2. Comparison of trends for BFI for all sites split by various classifications. Only sites with a significant ($\rho \le 0.05$) trend are included here as established by a Mann-Kendall test for monotonic trends across the full period of record. n is the number of sites, n_pos (n_neg) is the number of sites with positive (negative) trends, perc_pos (perc_neg) is the percentage of n with a positive (negative) trend. {#tbl-trends}

@@

: Table 3. Performance of model predictions for BFI for all sites split by various classifications. n is number of observatios, R^2^ is the coefficient of determination of a linear regression, MSE is mean-squared-error, RMSE is root-mean-squared-error, MAE is mean-absolute-error, NSE is Nash-Sucliffe efficiency, and pbias is percent bias. {#tbl-performance}

>>>>

: Basin-characteristic variables used as predictors in XGBoost model. {#tbl-predictors}

@@

The long-term average BFI for the 244 gauged reaches across Arizona is illustrated in @fig-instrumented-bfi . The highest BFI values (\>0.9) are found along the Grand Canyon in northwestern Arizona. The highly karstic geology of this region facilitates the rapid movement of subsurface flow to surface water and spring outlets [@chambless_deep-karst_2023]. Relatively high BFI values (\>0.8) are found at the spring-fed headwaters of the Verde River (Del Rio Spring) and the spring-fed headwaters of Fossil Creek. These results are consistent with interpolated BFI values reported by @wolock2003.

@@

![](images/BFITrends_significant.png)

@@

: Comparison of trends for BFI for all sites split by various classifications. Only sites with a significant ($\rho \le 0.05$) trend are included here as established by a Mann-Kendall test for monotonic trends across the full period of record. n is the number of sites, n_pos (n_neg) is the number of sites with positive (negative) trends, perc_pos (perc_neg) is the percentage of n with a positive (negative) trend. {#tbl-trends}

@@

: Performance of model predictions for BFI for all sites split by various classifications. n is number of observatios, R^2^ is the coefficient of determination of a linear regression, MSE is mean-squared-error, RMSE is root-mean-squared-error, MAE is mean-absolute-error, NSE is Nash-Sucliffe efficiency, and pbias is percent bias. {#tbl-performance}

Consistently negative pbias values across all classifications suggest an under-prediction of BFI.
